---
title: "APEC8211: Recitation 2"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = 'setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```


```{r, include = F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(gganimate)
library(ggtext)
library(gifski)
library(gganimate)
```

```{r, include = F, eval=F, cache = FALSE}
httpgd::hgd()
httpgd::hgd_browse()
```

# Outline

Review some concepts related to random variables

<!-- # main  -->
[1. CDF, PDF, PMF (Quick review)](#dist) 
  + [Exercise problem 1](#ex1)
  + [Exercise problem 2 (optional)](#ex2)

<!-- # To explain Jensen's inequality -->
[2. Mean and variance and covariance(Quick review)](#mean)
  + [Exercise problem 3](#ex3)
  + [Exercise problems 4 (optional)](#ex4)

[3. Introduction of Monte Calro Simulation](#monte)
[Supplement: Jensen's inequality (Quick review)](#jensen)

---
class: inverse, center, middle
name: dist

# CDF, PDF, and PMF

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
.content-box-red[**Distribution function**]
+ Cumulative distribution function (CDF)
+ Definition: <span style="color:red">The CDF of a random variable $X$ is $F(x) = Pr[X \leq x]$</span>
+ **Verbally**: CDF $F(x)$ tells us the probability of the event that random variable $X$ is <span style="color:red">less</span> than a value $x$. 

.left5[
```{r, echo=F}
# Define sequence of x-values  
x <-  seq(-4, 4, .01)
# CDF plot for a continuous variable
cdf_x <- pnorm(x)
# plot normal CDF
plot(
  x, cdf_x, type="l",
  xlab = "X",
  ylab = "Probability",
  main = "Continuous distribution function"
)
```
]
.right5[
```{r, dpi=36, echo=F}
# https://stackoverflow.com/questions/66266703/draw-discrete-cdf-in-r
x <- -1:5
prob <- ppois(q = x, lambda = 3)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(-1, max(x)), 
     ylim = c(0, 1),
     xlab = "X",
     ylab = "Probability",
     main = "Discrete distribution function"
)
points(x = x[-1], y = prob[-1], pch=19)
points(x = x[-1], y = prob[-n], pch=1)
for(i in 2:(n-1))
  points(x=x[i+0:1], y=prob[c(i,i)], type="l")
points(x=c(-2,x[2]), y=prob[c(1,1)], type="l")
points(x=c(x[n],6), y=prob[c(n,n)], type="l")
```
]

---
<!-- PDF and PMF (1) -->

.content-box-red[**Probability mass function (Discrete random variables)**]
+ **Definition**: $\color{red}{\pi(x) = Pr[X = x]}$
+ **Verbally**: The probability that $X$ equals the value $x$

<br>

.content-box-red[**Probability density function (Continuous random variables)**] 
+ **Definition**:  $\color{red}{f(x) = \frac{d}{dx}F(x)} \quad ( = \displaystyle \lim_{h\to\infty} \frac{F(x+h)-F(x)}{h})$
+ **Verbally**: Density function is defined as a very small change in the CDF (or the probability of the random variable falling within a particular range of values according to [wikipedia](https://en.wikipedia.org/wiki/Probability_density_function)). 

---
<!-- PDF and PMF (2) -->

.content-box-red[**Probability mass function (Discrete random variables)**]
+ **Definition**: $\color{red}{\pi(x) = Pr[X = x]}$
+ **Verbally**: The probability that $X$ equals the value $x$

<br>

.content-box-red[**Probability density function (Continuous random variables)**] 
+ **Definition**:  $\color{red}{f(x) = \frac{d}{dx}F(x)} \quad ( = \displaystyle \lim_{h\to\infty} \frac{F(x+h)-F(x)}{h})$
+ **Verbally**: Density function is a very small change in the CDF (or the probability of the random variable falling within a particular range of values according to [wikipedia](https://en.wikipedia.org/wiki/Probability_density_function)). 


<br>
.content-box-red[**Theorem 2.3: Properties of a PDF**] 

A function f(x) is a density function **if and only if** 

$$\begin{cases}
f(x) \ge 0 \text{ for all } x \\
\int_{-\infty}^\infty f(x)\,dx = 1
\end{cases}$$

+ You can use this condition to check whether a function is valid density function is or not!
<!-- if you are asked to show that a function f(x) is a valid density function, check whether f(x) satisfies these properties or not. -->


---
class: middle

.content-box-green[**Relationship between CDF and PDF**]
+ **From CDF to PDF**: $f(x) = \frac{d}{dx}F(x)$ </br> (by definition of PDF)

+ **From PDF to CDF**: $F(x) = Pr(X \leq x) = \int_{-\infty}^x f(t) dt$ </br> (as shown below)


```{r, child = 'prepareResults.rmd', cache=T}
```

---
name: ex1

# Exercise 1
.content-box-green[**Final Exam: 2021: Problem 1**] 

Define $\Phi(z)$ as the CDF of a standard normal random variable and $\phi(z)$ as its density function. 

(a) Write $Pr(Z \leq b)$ using $\Phi()$.

(b) Write $Pr(Z \leq b)$ as an integral. 

(c) Write $Pr(a \leq Z \leq b)$ using $\Phi()$.

(d) Write $Pr(a \leq Z \leq b)$ as an integral.

---
# Exercise 2
.content-box-green[**PSE Exercise 2.1**] 

Let $X \sim U[0,1]$. Find the PDF of random variable $Y=X^2$. 

---
class: inverse, center, middle
name: mean

# Mean and Variance and Covariance

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---

.content-box-red[**Mean and Variance**]

**Definition** 2.18, 2.19: 

+ The mean of $X$ is<span style='color:red'> $E[X]$</span>
+ The variance of $X$ is <span style='color:red'> $Var[X]=E[(X-E[X])^2]$</span> $= E[X^2] - (E[X])^2$


```{r,echo=F, cache=T}
# /*===========================================*/
#'=  Mean of X =
# /*===========================================*/
# /*===== Data generation =====*/
# --- Set the range of X --- #
x_left <- 2
x_right <- 10
x_center <- (x_left + x_right)/2

# --- Creating data --- #
x <- seq(x_left, x_right, length = 1000)
y <- dnorm(x, mean = x_center, sd = 1)
plot_data <- data.table(x = x, y = y)

# /*===== Visualization =====*/
plot_mean_X <- 
  ggplot(data = plot_data) +
    geom_line(aes(x = x, y = y), color = "black")+
    geom_vline(xintercept=x_center, color="red") +
    labs(y="Density", title="Density of X ~ N(6,1)") +
    annotate("text", x = 7, y = 0.01,
      label = "Mean E[X]=6",
      size = 3, color="red") +
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))
```

```{r, echo=F, cache=T}
# /*===========================================*/
#'= Variance =
# /*===========================================*/
# /*===== Data generation =====*/
# --- a list of the values of Var[X] --- #
ls_var <- c(1,4,9)
# --- Create an object to save results--- #
ls_res <- vector(mode='list', length=length(ls_var))
# --- Range of X --- #
x_left <- 2
x_right <- 10
x_center <- (x_left + x_right)/2

# --- Generate data by the value in ls_var --- #
for (i in seq(1:length(ls_var))){
  var <- ls_var[i]
  x <- seq(x_left, x_right, length = 1000)
  y <- dnorm(x, mean = x_center, sd = sqrt(var))
  ls_res[[i]] <- data.table(var = var, x = x, y = y)
}
res_total <- rbindlist(ls_res)

# /*===== Visualization =====*/
plot_var_X <- 
  ggplot(data = res_total) +
  geom_line(aes(x = x, y = y, color = interaction(var)))+
  labs(y = "Density", title = expression("Density of X ~ N(6," ~ sigma^2 ~ ") with various " ~ sigma^2)) +
  guides(color = guide_legend(title= expression(sigma^2 ~ "=")))+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


.left5[
```{r, dpi=36, echo=F, out.width = "100%", cache=T}
plot_mean_X
```
  ]

.right5[
```{r, dpi=36, echo=F, out.width = "100%", cache=T}
plot_var_X
```
  ]

---

.content-box-red[**Basic: Covariance**]

**Definition**
$$
\color{red}{Cov(X, Y) = E[(X-E[X])((Y-E[Y]))]}
$$

**Verbally**: Covariance measure the joint variability of two random variables. 

+ .content-box-green[Visualization]

```{r, dpi=36,, echo=F, fig.dim = c(12, 3)}
# /*===== Data generation =====*/
ls_a <- c(-0.8, -0.5, 0, 0.5, 0.8)
n = 1000
ls_result <- list()

set.seed(123)
x <- rnorm(n, 3, sd=2)
z <- rnorm(n, 3, sd=2)

for (i in seq(1:length(ls_a))){  
  a <- ls_a[i]
  y <- a*x + abs(1-abs(a))*z
  ls_result[[i]] <- data.table(a = a, x = x, y = y)
}

results <- 
  rbindlist(ls_result) %>%
  .[,type := paste0("Cov(X,Y)=",round(cov(x,y), digits = 1)), by=a]

# /*===== Visualization =====*/
ggplot(data = results)+
  geom_point(aes(x = x, y = y), size = 0.5)+
  facet_wrap(~ factor(type, unique(results$type)), scale = "free", nrow = 1)+
  labs(x = "X", y = "Y", title = "Plots of random variables (X,Y) with different sizes of covariances")+
  theme_bw()+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

.content-box-red[**Correlation**]

**Definition**
$$
\color{red}{Corr(X, Y) = \frac{Cov(X,Y)}{\sqrt{Var[X] Var[Y]}}}
$$

---
## Play around with data
.content-box-green[**Goal**]
+ Know some basic R functions 
+ See covariance is influenced by the change in scale but correlation is not. 

```{r, echo=T, eval=F}
# === Data === #
data(airquality)
# ?airquality

#/*--------------------------------*/
#' ## Basic functions of R
#/*--------------------------------*/
# --- histogram of Temperature (Temp) --- #
hist(airquality$Temp)
# frequency table can be obtained by running table(airquality$Temp)

# --- Mean of Temp (degrees F)--- #
mean(airquality$Temp)

# --- Variance of Temp --- #
var(airquality$Temp)
# sd(airquality$Temp) for standard deviation
# --- summary statistics of Temp --- #
summary(airquality$Temp)

#/*------------------------------------------*/
#' ## Relationship between Wind and Temp
#/*------------------------------------------*/
plot(airquality$Wind, airquality$Temp)

# === Covariance === #
cov(airquality$Wind, airquality$Temp)

# What happens if you change the unit of wind from mph to kmph (1mph=1.6kmph)
cov(airquality$Wind*1.6, airquality$Temp)

# === Correlation === #
cor(airquality$Wind, airquality$Temp)

# What happens if you change the unit of wind from mph to kmph (1mph=1.6kmph)
cor(airquality$Wind*1.6, airquality$Temp)
```

---


## E[ ], Var[ ], and Cov[ ] as operators

.content-box-red[**Expectation: E[ ]**]

<span style="color:red"> $E[ \,]$ is a linear operator (Linearity of expectation)</span>

For any constants $a$ and $b$, 

$$E[a+bX] = a + bE[X]$$


.content-box-red[**Variance: Var[ ]**]

$Var[ \,]$ is <span style="color:red">not</span> a linear operator

$$
Var[a+bX] = b^2E[X]
$$

because 
$$\begin{align*}
Var[a+bX]
  &= E[(a+bX - E[a+bX])^2] \\ 
  &= E[(a+bX - a-bE[X])^2] \\ 
  &= E[(b(X - E[X]))^2] \\ 
  &= E[b^2(X - E[X])^2] \\ 
  &= b^2 (X - E[X])^2 \\ 
  &= b^2 Var[X]
\end{align*}$$


---

.content-box-green[**Exercise 3: Lecture note 3, p14**] 

Prove these for continuous $(X,Y)$ with finite variances.

1. If $E[X]=0$ or $E[Y]=0$, $Cov(X,Y)=E[XY]$

2. If $X \perp\!\!\!\perp Y$, corr(X,Y)=0

3. $Var[X+Y] = Var[X] + Var[Y] + 2Cov(X,Y)$

4. If $X$ and $Y$ are correlated, $Var[X+Y] = Var[X] + Var[Y]$

---

.content-box-green[**Exercise 3: Final Exam: 2021: Problem 3**] 

The chi-squared distribution with $k$ degrees of freedom, denoted $\chi^2(k)$, is the distribution of $\sum_{i=1}^k Z^2_{i}$ and the $Z_i$ are independent $(Z_i \perp\!\!\!\perp Z_j)$. *You do not need to work with the CDF or density of a $\chi^2$ distribution to answer this question!*.

<br>

(a) Show that if $X$ is distributed $\chi^2(k)$ then $E[X]=k$.

<br>

(b) More work with expectation Let $K=Z^2_{1} + Z^2_{2}$, where $Z_j \sim N(0,1)$. Then $K \sim \chi^2(2)$. Another fact is that if $Z \sim N(0,1)$, then $E[Z^4]=3$. Use that fact to show that $Var[K]=4$. [Hint: $E[Z_j^4]$ is closely related to $Var[Z_j^2]$.]


---

class: inverse, center, middle
name: intro


# Monte Calro Simulation

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---

# Example: Binomial distribution


```{r, echo=F, eval=F}
p <- 1/2

# flip coins five times
# what is the expected numner of heads
# theoretically, it is E[X]=np=5*1/2=2.5

B <- 1000 # the number of iterations
n <- 5 # the number of trials

storage <- numeric(B)

for(i in 1:n){
  x <- sample(c(0,1), size=5, replace=TRUE)
  storage[[i]] <- sum(x)
}

hist(storage)

mean(storage)
```









---
class: inverse, center, middle
name: jensen

# Jensen's inequality

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---


## Motivation

Linearity of expectation cannot be used when the function inside $E[\,]$ is a nonlinear function.

.content-box-green[**Example:**] 
+ If $g(X)$ is a linear function (e.g., $g(x)=ax+b$)
  * $E[g(X)]=g(E[X]])$
+ If $g(X)$ is a nonlinear function (e.g., $g(x)=x^2$)
  * $E[g(X)] \neq g(E[X])$

---

(This is not the proof)

In the previous slide, we saw $Var[X]=E[(X-E[X])^2]=E[X^2] - (E[X])^2$. Because, $Var[X] \ge 0$ (by the way, $Var[X]=0$ if and only if $X$ is degenerate). 


$$E[X^2] - (E[X])^2 \ge 0$$
<p style="text-align: center;">or</p>
$$(E[X])^2 \leq E[X^2]$$
Define $g(x)=x^2$. Then it is written as
$$g(E[X]) \leq E[g(X)]$$

Generally, 
$$\begin{align*}
g(E[X]) \leq E[g(X)] \quad &\text{if } g(x) \text{ is a convex function} \\
E[g(X)] \leq g(E[X]) \quad &\text{if } g(x) \text{ is a concave function}
\end{align*}$$

---
.content-box-green[**Visualization**]

.panelset[ 

.panel[.panel-name[Example 1 : g(x) is convex]
.left5[

Suppose that $g(x)=x^2$.

```{r, echo=T}
set.seed(356)

# Create a sequence of X from a uniformal distribution
x <- runif(1000, 0, 10)

# /*===== Convex case: g(X)=X^2 =====*/
y <- x^2

figure_ex1 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+1, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- Add horizontal line for --- #
  geom_hline(yintercept = mean(y), color="blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)+5,
    label = paste0("E[g(X)]=", round(mean(y), 1)), size = 3, color = "blue") +
  # --- Add horizontal line for g(E[X]) --- #
  geom_hline(yintercept = mean(x)^2, color="darkgreen", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^2-5,
    label = paste0("g(E(X))=", round(mean(x)^2, 1)), size = 3, color = "darkgreen") +
  theme_bw()
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex1
```

$$\color{darkgreen}{g(E[X])} \leq \color{blue}{E[g(X)]}$$
  ]
]

.panel[.panel-name[Example 2: g(x) is concave]
.left5[

Suppose that $g(x)=\sqrt{x}$.

```{r, echo=T, out.width = "50%"}
# /*===== Convex case: g(X)=X^(1/2) =====*/
y <- x^(1/2)

figure_ex2 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+0.8, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- E[g(X)] --- #
  geom_hline(yintercept = mean(y), color = "blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)-0.2,
    label = paste0("E[g(X)]=", round(mean(y), 2)), size = 3, color = "blue") +
  # --- g(E[X]) --- #
  geom_hline(yintercept = mean(x)^(1/2), color = "darkgreen", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^(1/2)+0.2,
    label = paste0("g(E(X))=", round(mean(x)^(1/2), 2)), size = 3, color = "darkgreen") +
  theme_bw()
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex2
```
$$\color{blue}{E[g(X)]} \leq \color{darkgreen}{g(E[X])}$$
  ]
]
]


---
<!-- https://stackoverflow.com/questions/53481699/customize-font-size-for-all-the-slides-in-xaringan -->
<!-- class: my-one-page-font -->

# Solution to Exercise 2 (slide 1/2)

.content-box-green[**Idea:**]

We want to derive the PDF of $Y$ which is defined by $Y=X^2$. Since $0 \leq X \leq 1$, $0 \leq Y \leq 1$. Recall the definition of PDF. Let $G(y)$ be the CDF of Y and $g(y)$ be the PDF of Y. Then, $g(y)=\frac{d}{dy}G(y)$ by the definition of PDF. That is, once you get $G(y)$, you can derive $g(y)$. So, let's start with the CDF of $Y$. 

.content-box-green[**Solution:**]

By the definition of CDF, 
$$G(y) = Pr(Y \leq y).$$
Substituting $Y=X^2$, 
$$\begin{align*}
G(y) 
&= Pr(X^2 \leq y)\\
&= Pr(0 \leq X \leq \sqrt{y})
\end{align*}$$
(The last equality is the combination of the fact that $X^2 \leq y \iff -\sqrt{y} \leq X \leq \sqrt{y}$ and the condition $X \ge 0$.)

---

# Solution to Exercise 2 (slide 1/2)

Since we know that $X$ is uniformly distributed ($\(X \sim U[0,1]\)$), the PDF of $X$ is $1$ for $0 \leq X \leq 1$. Also, note that the range $[0, \sqrt{y}]$ is contained in the range of the PDF of $X$.

$$
G(y) = Pr(0 \leq X \leq \sqrt{y}) = \int_{0}^{\sqrt{y}} 1 dx = \sqrt{y}.
$$

Thus, PDF of $Y$ is 
$$\begin{align*}
g(y) = \frac{d}{dy}G(y) = \frac{1}{2\sqrt{y}} \quad (0 \leq y \leq 1)
\end{align*}$$


---
Question:


The problem sets in the assignment covers the materials from the previous week. 


**Question: Should I cover the material we learned from this week,  the previous week or the mix of the two?**

<br>

**option (1): lab covers the material from this week**

.content-box-green[**pro:**] The material in the lab is the review of the class on this week and it will be helpful for the next assignment. 

.content-box-red[**con:**] You may not have time to review the textbook or class-notes, so the lab would be boring. 

<br>

**option (2): lab covers the material from the previous week**
  
.content-box-green[**pro:**] You have enough time to review the textbook and class-note, so it is more likely to understand the lab

.content-box-red[**con:**]





