---
title: "Recitation 2"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = 'setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```


```{r, include = F, eval=F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)

httpgd::hgd()
httpgd::hgd_browse()
```


```{r}
figure_theme <- 
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )


```

# Outline

1. [Review some concepts related to random variables](#intro)
  * CDF, PDF, PMF
  * Mean and variance, covariance
3. [Exercise](#Exercise)

---
# CDF, PDF, PMF, Quantile function
.content-box-red[**CDF**]
+ Definition: The CDF of a random variable X is $F(x) = Pr[X \leq x]$
+ **Verbally**: CDF $F(x)$ specifies the probability of the event that random variable $X$ is less than $x$. 

```{r, echo=F}
# visualizing CDF for discrete and continuous case. 
# Define sequence of x-values
x <- seq(-4, 4, .01)
# #calculate normal CDF probabilities
# cdf <- pnorm(x)
# pdf <- rnorm(x)
# #plot normal CDF
# plot(x, cdf, type="l")
# plot(x, pdf, type="l")
```
---
.content-box-red[**Probability mass function (Discrete random variables)**]
+ Definition: $\pi(x) = Pr[X = x]$

.content-box-red[**Probability density function (Continuous random variables)**] 
+ Definition: $f(x) = \frac{d}{dx}F(x)$


---
# Properties of PDF (and PMF)



---
# Mean, Variance and Covariance:
.content-box-red[**Mean: E[X]**]
+ Definition: <span style='color:red'>The mean of $X$ is $E[X]$</span>
+ How to calculate it?:
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, echo=F, out.width = "50%"}
x_left = 2
x_right = 10
x_center = (x_left+x_right)/2
x <- seq(x_left, x_right, length = 1000)
y <- dnorm(x, mean = x_center, sd = 1)
plot_data <- data.table(x = x, y = y)

ggplot(data = plot_data) +
  geom_line(aes(x = x, y = y), color = "black")+
  labs(y="Density", title="Density of X ~ N(6,1)") +
  geom_vline(xintercept=x_center, color="red") +
  annotate("text",
    x = 7, y = 0.01,
    label = "Mean E[X]=6",
    size = 3, color="red"
  ) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Mean, Variance and Covariance:
.content-box-red[**Variance: Var[X]**]
+ Definition: <span style='color:red'>The variance of $X$ is $Var[X]=E[(X-E[X])^2]$</span>
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, echo=F, out.width = "50%"}
ls_res <- list()
ls_var <- c(1,4,9)

for (i in seq(1:length(ls_var))){
  var <- ls_var[i]
  x <- seq(x_left, x_right, length = 1000)
  y <- dnorm(x, mean = x_center, sd = sqrt(var))
  ls_res[[i]] <- data.table(var = var, x = x, y = y)
}

res_total <- rbindlist(ls_res)

ggplot(data = res_total) +
  geom_line(aes(x = x, y = y, color=interaction(var)))+
  labs(y="Density", title=expression("Density of X ~ N(6," ~ sigma^2 ~ ") with various " ~ sigma^2)) +
  guides(color=guide_legend(title= expression(sigma^2 ~ "=")))+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Mean, Variance and Covariance:
.content-box-red[**Coariance: Cov[X, Y]**]
+ Definition: <span style='color:red'>The covariance between X and Y are</span>

$$
\color{red}{Cov(X, Y) = E[(X-E[X])((Y-E[Y]))]}
$$
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, echo=F, fig.dim = c(12, 3)}
list_a <- c(-0.8, -0.5, 0, 0.5, 0.8)
n = 1000
ls_res <- list()

for (i in seq(1:length(list_a))){
  set.seed(123)
  a <- list_a[i]
  x <- rnorm(n, 3, sd=2)
  z <- rnorm(n, 3, sd=2)
  y <- a*x + abs(1-abs(a))*z
  ls_res[[i]] <- data.table(a = a, x = x, y = y)
}

res_total <- 
  rbindlist(ls_res) %>%
  .[,order := .GRP,by=a] %>%
  .[,type := paste0("Cov(X,Y)=",round(cov(x,y), digits = 1)), by=a]

ggplot(data = res_total)+
  geom_point(aes(x = x, y = y), size=0.5)+
  facet_wrap(~ factor(type, unique(res_total$type)), scale="free", nrow = 1)+
  labs(x="X", y="Y", title="Plots of random variables (X,Y) with different sizes of covariances")+
  theme_bw()+
  theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

---
class: middle

# Expectation, Variance and Covariance as operators
In Econometrics analysis, we often use $E[\,]$, $Var[\,]$, $Cov[\,]$ as operators (i.e., functions). The following are useful knowledge that you should know. 

---
class: middle

.content-box-red[**Expectation**] 

Expectation is a linear operator.

That is, for any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $E[a+bX]=a+bE[X]$

(2) $E[X+Y]=E[X]+E[Y]$

+ Note: Linearity of expectation only applies to the sum of random variables (random variables are in the linear function) 
<!-- + A function $g(x)$ called a linear operator if (1) $f(x+y)=f(x)+f(y)$, and (2) $g(cx)=cg(x)$ for all $x$ and constant $c$.  -->

.content-box-green[**Question**]
+ True or False: $E[X+X^2]=E[X]+E[X^2]$?
+ True or False: $E[XY]=E[X]E[Y]$

---
class: middle

.content-box-red[**Variance**]

Variance is not a linear operator except for special case. 

For any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $Var[X]=E[X^2] - (E[X])^2 \quad (\text{Simply, another definition of } Var[X]$ )

(2) $Var[aX]=a^2Var[X]$

(3) $Var[a+bX]=b^2Var[X]$

(4) $Var[X+Y]=Var[X]+Var[Y] + Cov(X, Y)$

.content-box-green[**Question**]
+ In what condition does $Var[X+Y]=Var[X]+Var[Y]$ hold?


.content-box-green[**Exercise**]
+ Let's prove (1) and (2) 

---
class: middle

.content-box-red[**Covariance**]

For any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $Cov[X, Y]=E[XY]-E[X]E[Y] \quad (\text{Simply, another definition of } Cov[X, Y]$)

(2) $Cov[aX, Y]=aCov[X,Y]$

(3) $Cov[X, a+bY]=bCov[X,Y]$

.content-box-green[**Question**]
+ In what condition does $Cov[X, Y]=0$ hold?

.content-box-green[**Exercise**]
+ Let's prove (1)-(3)

---

# Jensen's inequality
In the previous slide, we saw $Var[X]=E[(X-E[X])^2]=E[X^2] - (E[X])^2$. So, $Var[X] \ge 0$ (by the way, $Var[X]=0$ if and only if $X$ is degenerate). 

So, 
$$E[X^2] - (E[X])^2 \ge 0$$
<p style="text-align: center;">or</p>
$$(E[X])^2 \leq E[X^2]$$
Define $g(x)=x^2$. The it is written as
$$g(E[X]) \leq E[g(X)]$$

Generally, 
$$\begin{align*}
g(E[X]) \leq E[g(X)] \quad &\text{if } g(x) \text{ is a convex function} \\
E[g(X)] \leq g(E[X]) \quad &\text{if } g(x) \text{ is a concave function}
\end{align*}$$

---
.content-box-green[**Implication**]

.panelset[ 

.panel[.panel-name[Example 1 : g(x) is convex]
.left5[
```{r, echo=T}
set.seed(356)
x <- runif(1000, 0, 10)

# /*===== Convex case: g(x)=x^2 =====*/
y <- x^2

figure_ex1 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+1, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- E[g(X)] --- #
  geom_hline(yintercept = mean(y), color="blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)+5,
    label = paste0("E[g(X)]=", round(mean(y), 1)), size = 3, color = "blue") +
  # --- g(E[X]) --- #
  geom_hline(yintercept = mean(x)^2, color="green", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^2-5,
    label = paste0("g(E(X))=", round(mean(x)^2, 1)), size = 3, color = "green") +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex1
```

$$\color{lightgreen}{g(E[X])} \leq \color{blue}{E[g(X)]}$$
  ]
]

.panel[.panel-name[Example 2: g(x) is concave]
.left5[
```{r, echo=T, out.width = "50%"}
# /*===== Convex case: g(x)=x^(1/2) =====*/
y <- x^(1/2)

figure_ex2 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+0.8, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- E[g(X)] --- #
  geom_hline(yintercept = mean(y), color = "blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)-0.2,
    label = paste0("E[g(X)]=", round(mean(y), 2)), size = 3, color = "blue") +
  # --- g(E[X]) --- #
  geom_hline(yintercept = mean(x)^(1/2), color = "green", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^(1/2)+0.2,
    label = paste0("g(E(X))=", round(mean(x)^(1/2), 2)), size = 3, color = "green") +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex2
```
$$\color{blue}{E[g(X)]} \leq \color{lightgreen}{g(E[X])}$$
  ]
]
]

