---
title: "APEC8211: Recitation 2"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = 'setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```


```{r, include = F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(gganimate)
library(ggtext)
library(gifski)
library(gganimate)
```

```{r, include = F, eval=F, cache = FALSE}
httpgd::hgd()
httpgd::hgd_browse()
```

# Outline

Review some concepts related to random variables

<!-- # main  -->
[1. CDF, PDF, PMF (Quick review)](#dist) 
  + [Exercise problems](#ex1)

<!-- # To explain Jensen's inequality -->
[2. Mean and variance (Quick review)](#mean)
  + [Exercise problems](#ex2)

[3. Jensen's inequality (Quick review)](#jensen)

---
class: inverse, center, middle
name: dist

# CDF, PDF, and PMF

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
.content-box-red[**Distribution function**]
+ Cumulative distribution function (CDF)
+ Definition: <span style="color:red">The CDF of a random variable $X$ is $F(x) = Pr[X \leq x]$</span>
+ **Verbally**: CDF $F(x)$ tells us the probability of the event that random variable $X$ is less than a value $x$. 

.left5[
```{r, echo=F}
# Define sequence of x-values  
x <-  seq(-4, 4, .01)
# CDF plot for a continuous variable
cdf_x <- pnorm(x)
# plot normal CDF
plot(
  x, cdf_x, type="l",
  xlab = "X",
  ylab = "Probability",
  main = "Continuous distribution function"
)
```
]
.right5[
```{r, dpi=36, echo=F}
# https://stackoverflow.com/questions/66266703/draw-discrete-cdf-in-r
x <- -1:5
prob <- ppois(q = x, lambda = 3)
n <- length(x)

plot(x = NA, y = NA, pch = NA, 
     xlim = c(-1, max(x)), 
     ylim = c(0, 1),
     xlab = "X",
     ylab = "Probability",
     main = "Discrete distribution function"
)
points(x = x[-1], y = prob[-1], pch=19)
points(x = x[-1], y = prob[-n], pch=1)
for(i in 2:(n-1))
  points(x=x[i+0:1], y=prob[c(i,i)], type="l")
points(x=c(-2,x[2]), y=prob[c(1,1)], type="l")
points(x=c(x[n],6), y=prob[c(n,n)], type="l")
```
]

---
<!-- PDF and PMF (1) -->

.content-box-red[**Probability mass function (Discrete random variables)**]
+ **Definition**: $\color{red}{\pi(x) = Pr[X = x]}$
+ **Verbally**: The probability that $X$ equals the value $x$

<br>

.content-box-red[**Probability density function (Continuous random variables)**] 
+ **Definition**:  $\color{red}{f(x) = \frac{d}{dx}F(x)} \quad ( = \displaystyle \lim_{h\to\infty} \frac{F(x+h)-F(x)}{h})$
+ **Verbally**: Density function is a very small change in the CDF (or the probability of the random variable falling within a particular range of values according to [wikipedia](https://en.wikipedia.org/wiki/Probability_density_function)). 

---
<!-- PDF and PMF (2) -->

.content-box-red[**Probability mass function (Discrete random variables)**]
+ **Definition**: $\color{red}{\pi(x) = Pr[X = x]}$
+ **Verbally**: The probability that $X$ equals the value $x$

<br>

.content-box-red[**Probability density function (Continuous random variables)**] 
+ **Definition**:  $\color{red}{f(x) = \frac{d}{dx}F(x)} \quad ( = \displaystyle \lim_{h\to\infty} \frac{F(x+h)-F(x)}{h})$
+ **Verbally**: Density function is a very small change in the CDF (or the probability of the random variable falling within a particular range of values according to [wikipedia](https://en.wikipedia.org/wiki/Probability_density_function)). 


<br>
.content-box-red[**Theorem 2.3: Properties of a PDF**] 

A function f(x) is a density function **if and only if** 

$$\begin{cases}
f(x) \ge 0 \text{ for all } x \\
\int_{-\infty}^\infty f(x)\,dx = 1
\end{cases}$$

<!-- if you are asked to show that a function f(x) is a valid density function, check whether f(x) satisfies these properties or not. -->


---
class: middle

.content-box-green[**Relationship between CDF and PDF**]
+ **From CDF to PDF**: $f(x) = \frac{d}{dx}F(x)$ </br> (by definition of PDF)

+ **From PDF to CDF**: $F(x) = Pr(X \leq x) = \int_{-\infty}^x f(t) dt$ </br> (as shown below)


```{r, child = 'prepareResults.rmd', cache=T}
```

---
name: ex1

# Exercise 1
.content-box-green[**Final Exam: 2021: Problem 1**] 

Define $\Phi(z)$ as the CDF of a standard normal random variable and $\phi(z)$ as its density function. 

(a) Write $Pr(Z \leq b)$ usinzg $\Phi()$.

(b) Write $Pr(Z \leq b)$ as an integral. 

(c) Write $Pr(a \leq Z \leq b)$ using $\Phi()$.

(d) Write $Pr(a \leq Z \leq b)$ as an integral.

---
# Exercise 2
.content-box-green[**PSE Exercise 2.1**] 

Let $X \sim U[0,1]$. Find the PDF of random variable $Y=X^2$. 

---
class: inverse, center, middle
name: mean

# Mean and Variance and their properties

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---
<!-- mean and variance 1 -->
**Definition** 2.18, 2.19: 
+ The mean of $X$ is<span style='color:red'> $E[X]$</span>
+ The variance of $X$ is <span style='color:red'> $Var[X]=E[(X-E[X])^2]$</span>

---
<!-- mean and variance 2 -->
**Definition** 2.18, 2.19: 
+ The mean of $X$ is<span style='color:red'> $E[X]$</span>
+ The variance of $X$ is <span style='color:red'> $Var[X]=E[(X-E[X])^2]$</span>

.content-box-green[Visualization]

.panelset[ 

.panel[.panel-name[Code 1: Mean]
```{r,echo=T, cache=T}
# /*===========================================*/
#'=  Mean of X =
# /*===========================================*/
# /*===== Data generation =====*/
# --- Set the range of X --- #
x_left <- 2
x_right <- 10
x_center <- (x_left + x_right)/2

# --- Creating data --- #
x <- seq(x_left, x_right, length = 1000)
y <- dnorm(x, mean = x_center, sd = 1)
plot_data <- data.table(x = x, y = y)

# /*===== Visualization =====*/
plot_mean_X <- 
  ggplot(data = plot_data) +
    geom_line(aes(x = x, y = y), color = "black")+
    labs(y="Density", title="Density of X ~ N(6,1)") +
    geom_vline(xintercept=x_center, color="red") +
    annotate("text", x = 7, y = 0.01,
      label = "Mean E[X]=6",
      size = 3, color="red") +
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))
```
  ]

.panel[.panel-name[Code 2: Variance]
```{r, echo=T, cache=T}
# /*===========================================*/
#'= Variance =
# /*===========================================*/
# /*===== Data generation =====*/
# --- a list of the values of Var[X] --- #
ls_var <- c(1,4,9)
# --- Create an object to save results--- #
ls_res <- vector(mode='list', length=length(ls_var))
# --- Range of X --- #
x_left <- 2
x_right <- 10
x_center <- (x_left + x_right)/2

# --- Generate data by the value in ls_var --- #
for (i in seq(1:length(ls_var))){
  var <- ls_var[i]
  x <- seq(x_left, x_right, length = 1000)
  y <- dnorm(x, mean = x_center, sd = sqrt(var))
  ls_res[[i]] <- data.table(var = var, x = x, y = y)
}
res_total <- rbindlist(ls_res)

# /*===== Visualization =====*/
plot_var_X <- 
  ggplot(data = res_total) +
  geom_line(aes(x = x, y = y, color = interaction(var)))+
  labs(y = "Density", title = expression("Density of X ~ N(6," ~ sigma^2 ~ ") with various " ~ sigma^2)) +
  guides(color = guide_legend(title= expression(sigma^2 ~ "=")))+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```
  ]

.panel[.panel-name[Graphs]

.left5[
```{r, dpi=36, echo=F, out.width = "100%", cache=T}
plot_mean_X
```
  ]

.right5[
```{r, dpi=36, echo=F, out.width = "100%", cache=T}
plot_var_X
```
  ]
  ]
]

---
## E[] and Var[] as operators

.content-box-red[**Expectation**]

<span style="color:red"> $E[ \,]$ is a linear operator (Linearity of expectation)</span>

For any constants $a$ and $b$, 

$$E[a+bX] = a + bE[X]$$

---
## E[ ] and Var[ ] as operators

.content-box-red[**Expectation: E[]**]

<span style="color:red"> $E[ \, ]$ is a linear operator (i.e. Linearity of expectation)</span>

For any constants $a$ and $b$, 

$$E[a+bX] = a + bE[X]$$

.content-box-red[**Variance: Var[]**]
$$\begin{align*}
\text{(1) } &Var[X] = E[X^2] - (E[X])^2 \\
\text{(2) } &Var[a+bX] = b^2E[X]
\end{align*}$$

.content-box-green[**Question**]

Prove (1) and (2) using linearity of expectation. 

---

name: jensen

# Jensen's inequality

(This is not the proof)

In the previous slide, we saw $Var[X]=E[(X-E[X])^2]=E[X^2] - (E[X])^2$. So, $Var[X] \ge 0$ (by the way, $Var[X]=0$ if and only if $X$ is degenerate). 

So, 
$$E[X^2] - (E[X])^2 \ge 0$$
<p style="text-align: center;">or</p>
$$(E[X])^2 \leq E[X^2]$$
Define $g(x)=x^2$. The it is written as
$$g(E[X]) \leq E[g(X)]$$

Generally, 
$$\begin{align*}
g(E[X]) \leq E[g(X)] \quad &\text{if } g(x) \text{ is a convex function} \\
E[g(X)] \leq g(E[X]) \quad &\text{if } g(x) \text{ is a concave function}
\end{align*}$$

---
.content-box-green[**Visualization**]

.panelset[ 

.panel[.panel-name[Example 1 : g(x) is convex]
.left5[

Suppose that $g(x)=X^2$.

```{r, echo=T}
set.seed(356)
x <- runif(1000, 0, 10)

# /*===== Convex case: g(X)=X^2 =====*/
y <- x^2

figure_ex1 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+1, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- E[g(X)] --- #
  geom_hline(yintercept = mean(y), color="blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)+5,
    label = paste0("E[g(X)]=", round(mean(y), 1)), size = 3, color = "blue") +
  # --- g(E[X]) --- #
  geom_hline(yintercept = mean(x)^2, color="darkgreen", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^2-5,
    label = paste0("g(E(X))=", round(mean(x)^2, 1)), size = 3, color = "darkgreen") +
  theme_bw()
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex1
```

$$\color{darkgreen}{g(E[X])} \leq \color{blue}{E[g(X)]}$$
  ]
]

.panel[.panel-name[Example 2: g(x) is concave]
.left5[

Suppose that $g(x)=\sqrt{x}$.

```{r, echo=T, out.width = "50%"}
# /*===== Convex case: g(X)=X^(1/2) =====*/
y <- x^(1/2)

figure_ex2 <-
  ggplot()+
  geom_point(aes(x = x, y = y))+
  # --- E[X] --- #
  geom_vline(xintercept = mean(x), color = "red", linetype = "dashed")+
  annotate("text", x = mean(x)+0.8, y = 0.01,
    label = paste0("E[X]=", round(mean(x), 1)), size = 3, color = "red") +
  # --- E[g(X)] --- #
  geom_hline(yintercept = mean(y), color = "blue", linetype = "dashed")+
  annotate("text", x = 1, y = mean(y)-0.2,
    label = paste0("E[g(X)]=", round(mean(y), 2)), size = 3, color = "blue") +
  # --- g(E[X]) --- #
  geom_hline(yintercept = mean(x)^(1/2), color = "darkgreen", linetype = "dashed")+
  annotate("text", x = 1, y = mean(x)^(1/2)+0.2,
    label = paste0("g(E(X))=", round(mean(x)^(1/2), 2)), size = 3, color = "darkgreen") +
  theme_bw()
```
  ]
.right5[
```{r, echo=F, out.width = "100%"}
figure_ex2
```
$$\color{blue}{E[g(X)]} \leq \color{darkgreen}{g(E[X])}$$
  ]
]
]

---
<!-- https://stackoverflow.com/questions/53481699/customize-font-size-for-all-the-slides-in-xaringan -->
<!-- class: my-one-page-font -->

# Solution to Exercise 2
We want to derive the PDF of $Y=X^2$. Since $0 \leq X \leq 1$, $0 \leq Y \leq 1$. Let $G(y)$ be the CDF of Y and $g(y)$ be the PDF of Y. By definition of PDF, $g(y)=\frac{d}{dy}G(y)$ . So, once you get $G(y)$, you can derive $g(y)$. 

By definition of CDF, 
$$G(y) = Pr(Y \leq y).$$
Since $Y=X^2$, 
$$\begin{align*}
G(y) 
&= Pr(X^2 \leq y)\\
&= Pr(0 \leq X \leq \sqrt{y}) \quad (\because X \ge 0).\\
\end{align*}$$
We know that $X \sim U[0,1]$. So, 
$$
G(y) = \int_{0}^{\sqrt{y}} 1 dx = \sqrt{y}.
$$
Thus, 
$$\begin{align*}
g(y) = \frac{d}{dy}G(y) = \frac{1}{2\sqrt{y}} \quad (0 \leq y \leq 1)
\end{align*}$$

---





