---
title: "Recitation 2"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = 'setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```


```{r, include = F, eval=F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)

httpgd::hgd()
httpgd::hgd_browse()
```




# Outline

1. [Review some concepts related to random variables](#intro)
  * CDF, PDF, PMF
  * Mean and variance 
3. [Exercise](#Exercise)

---
# CDF, PDF, PMF, Quantile function
.content-box-red[**CDF**]
+ Definition: The CDF of a random variable X is $F(x) = Pr[X \leq x]$

```{r, echo=F}
# Define sequence of x-values
x <- seq(-4, 4, .01)
# #calculate normal CDF probabilities
# cdf <- pnorm(x)
# pdf <- rnorm(x)
# #plot normal CDF
# plot(x, cdf, type="l")
# plot(x, pdf, type="l")
```
---
.content-box-red[**Probability mass function (Discrete random variables)**]
+ Definition: $\pi(x) = Pr[X = x]$

.content-box-red[**Probability density function (Continuous random variables)**] 
+ Definition: $f(x) = \frac{d}{dx}F(x)$


---
# Properties of PDF (and PMF)



---
# Mean, Variance and Covariance:
.content-box-red[**Mean: E[X]**]
+ Definition: <span style='color:red'>The mean of $X$ is $E[X]$</span>
+ How to calculate it?:
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, echo=F, out.width = "50%"}
x_left = 2
x_right = 10
x_center = (x_left+x_right)/2
x <- seq(x_left, x_right, length = 1000)
y <- dnorm(x, mean = x_center, sd = 1)
plot_data <- data.table(x = x, y = y)

ggplot(data = plot_data) +
  geom_line(aes(x = x, y = y), color = "black")+
  labs(y="Density", title="Plot of X ~ N(6,1)") +
  geom_vline(xintercept=x_center, color="red") +
  annotate("text",
    x = 7, y = 0.01,
    label = "Mean E[X]=6",
    size = 3, color="red"
  ) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Mean, Variance and Covariance:
.content-box-red[**Variance: Var[X]**]
+ Definition: <span style='color:red'>The variance of $X$ is $Var[X]=E[(X-E[X])^2]$</span>
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, echo=F, out.width = "50%"}
ls_res <- list()
ls_var <- c(1,4,9)

for (i in seq(1:length(ls_var))){
  var <- ls_var[i]
  x <- seq(x_left, x_right, length = 1000)
  y <- dnorm(x, mean = x_center, sd = sqrt(var))
  ls_res[[i]] <- data.table(var = var, x = x, y = y)
}

res_total <- rbindlist(ls_res)

ggplot(data = res_total) +
  geom_line(aes(x = x, y = y, color=interaction(var)))+
  labs(y="Density") +
  guides(color=guide_legend(title="Var[X]="))+
  theme_bw()
```


---
.content-box-red[**Coariance: cov[X, Y]**]
+ Definition: <span style='color:red'>The covariance between X and Y are</span>

$$
\color{red}{cov(X, Y) = E[(X-E[X])((Y-E[Y]))]}
$$
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r}
set.seed(1213)

n=1000
x <- rnorm(n, 3, sd=2)
y <- rnorm(n, 3, sd=2)

plot_data <- data.table(x = x, y = y)

ggplot(data = plot_data) +
  geom_point(aes(x = x, y = y)) +
  labs(title = paste0("Cov(X,Y)=", round(cov(x,y), 1))) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))


x <- mvrnorm(n = 1, mu, Sigma, tol = 1e-6, empirical = FALSE, EISPACK = FALSE)



list_a <- c(-1, -0.5, 0, 0.5, 1)


ls_res <- list()

for (i in seq(1:length(list_a))){
  set.seed(123)
  a <- list_a[i]
  x <- rnorm(n, 3, sd=2)
  z <- rnorm(n, 3, sd=2)
  y <- a*x + abs(1-abs(a))*z
  ls_res[[i]] <- data.table(a = a, x = x, y = y)
}

res_total <- 
  rbindlist(ls_res)


ggplot(data = plot_data) +
  geom_point(aes(x = x, y = y)) +
  labs(title = paste0("Cov(X,Y)=", round(cov(x,y), 1))) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))


```



---
# Expectation, Variance and Covariance as operators
In Econometrics analysis, we use $E[\,]$, $Var[\,]$, $Cov[\,]$ as operators (i.e., functions). Bellows are useful properties that you should know. 

For any constants $a$ and $b$ and random variables $X$ and $Y$:

.content-box-red[**Expectation**] 
+ $E[a+bX]=a+bE[X]$
+ $E[aX+bY]=aE[X]+bE[Y]$
Note: Expectation is a linear operator.
+ A function $g(x)$ called a linear operator if (1)$f(x+y)=f(x)+f(y)$, and (2)$g(cx)=cg(x)$ for all $x$ and constant $c$. 
<br>

.content-box-red[**Variance**]
+ $Var[X]=E[X^2] - E[X]^2 \quad (\text{Simply, another definition of } Var[X]$ )
+ $Var[aX]=a^2Var[X]$
+ $Var[a+bX]=b^2Var[X]$
+ $Var[aX+bY]=a^2Var[X]+b^2Var[Y] + 2abCov(X, Y)$


---
.content-box-red[**Covariance**]
+ $Cov[X, Y]=E[XY]-E[X]E[Y] \quad (\text{Simply, another definition of } Cov[X, Y]$)
+ $Cov[aX, Y]=aCov[X,Y]$
+ $Cov[X, a+bY]=bCov[]$

---


class: inverse, center, middle
name: Exercise


# Exercise

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---

