---
title: "APEC8212: Recitation 13"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: ../xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = '../setup.rmd', cache = FALSE}
```

```{r xaringan-scribble, echo=FALSE}
xaringanExtra::use_scribble()
```


```{r setup, include=FALSE, cache = FALSE, eval=F}
options(htmltools.dir.version = FALSE)

# /*===== Reference =====*/
suppressMessages(library(RefManageR))

BibOptions(
  check.entries = FALSE,
  bib.style = "authoryear",
  style = "markdown",
  hyperlink = FALSE,
  dashed = TRUE,
  max.names = 2,
  longnamesfirst = FALSE
)
bib <- ReadBib("cite.bib")
```

```{r, include = F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(gganimate)
library(gifski)
library(gganimate)
# library(learnr)
```

```{r, include = F, eval=F, cache = FALSE}
httpgd::hgd()
httpgd::hgd_browse()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```


.content-box-green[**A Useful tip:**]

hitting letter "o" key will give you a panel view of the slides

---
class: middle

# Outline
1.[Review panel data analysis](#Resampling)

2.[Implementation with R](#R)

2. [Code review for assignment 7](#code)

???
+ Okay, today, we gonna do a quick review of resampling methods, and do one exercise problem. The exercise problem is from last year's final exam. 

---
class: inverse, center, middle
name: Resampling

# Panel data analysis

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>

---
class: middle

# Panel (longitudinal) Data

.content-box-green[**Definition**]

Data follows the same individuals, families, firms, cities, states or whatever, across time

<br>

.content-box-green[**Example**]

+ Randomly select people from a population at a given point in time
+ Then the same people are re-interviewed at several subsequent points in time, which would result in data on wages, hours, education, and so on, for the same group of people in different years.

???
+ As you know, panel data follows the same individuals...
+ With panel data, you can address the issues that cannot be studied using time series or cross-sectional data alone
  * For example, you can control unobserved individual heterogeneity  

---
class: middle

.content-box-green[**Central Question**]

Can we do anything to deal with endogeneity problem taking advantage of the panel data structure?


---
class: middle

# One-way error component model

$$Y_{it} = X_{it}^{\prime} \beta + \color{blue}{e_{it}}$$

<p style="text-align: center;">&darr;</p>

$$Y_{it} = X_{it}^{\prime} \beta + \color{blue}{u_{i} + \varepsilon_{it}}$$

+ $u_{i}$: a (time-invariant unobserved) individual-specific effect 
+ $\varepsilon_{it}$: idiosyncratic errors 


???

---
# Random effect model

$$Y_{it} = X_{it}^{\prime} \beta + \color{blue}{u_{i} + \varepsilon_{it}}$$

---
# Fixed effect model

[Mundlak (1961)](https://www.jstor.org/stable/1235460)

$$Y_{it} = X_{it}^{\prime} \beta + \color{blue}{u_{i} + \varepsilon_{it}}$$

+ $u_{i}$: an individual-specific effect 

.content-box-red[**Assumption**]



within-transformation or demeaning.

---
class: middle

Lots of interesting discussions on fixed effects estimation for causal inference are going on. For example, 

[Imai and Kim (2019)](https://onlinelibrary.wiley.com/doi/abs/10.1111/ajps.12417)

[Imai and Kim (2021)](https://www.cambridge.org/core/journals/political-analysis/article/abs/on-the-use-of-twoway-fixed-effects-regression-models-for-causal-inference-with-panel-data/F10006D0210407C5F9C7CAC1EEE3EF0D)

[Millimet and Bellemare (2023)](https://docs.iza.org/dp16202.pdf)

In the context of climate economics, 

[Blanc and Schlenker (2017)](https://www.journals.uchicago.edu/doi/abs/10.1093/reep/rex016?journalCode=reep)

???
+ Okay, these are the very basic knowledge about fixed effect estimations, but 
+ Causal inference is all about the question of how to credibly estimate the counterfactual outcomes through the comparison of treated and control observations



---

.content-box-green[**Summary**]

+ As long as the omitted variable that affects both the dependent and independent variables are constant over time (time-invariant), then using only the variations over time (ignoring variations across cross-sectional units) can eliminate the omitted variable bias

+ First-differencing the data and then regressing changes on changes does the trick of ignoring variations across cross-sectional units

+ Of course, first-differencing is possible only because the same cross-sectional units are observed multiple times over time.

---
class: inverse, center, middle
name: intro

# RE and FE estimation with R

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---




---




