---
title: "APEC8211: Recitation 6"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: ../xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = '../setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```

```{r setup, include=FALSE, cache = FALSE}
options(htmltools.dir.version = FALSE)

# /*===== Reference =====*/
suppressMessages(library(RefManageR))

BibOptions(
  check.entries = FALSE,
  bib.style = "authoryear",
  style = "markdown",
  hyperlink = FALSE,
  dashed = TRUE,
  max.names = 2,
  longnamesfirst = FALSE
)
bib <- ReadBib("cite.bib")
```

```{r, include = F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(gganimate)
library(gifski)
library(gganimate)
# library(learnr)
```

```{r, include = F, eval=F, cache = FALSE}
httpgd::hgd()
httpgd::hgd_browse()
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```


.content-box-green[**A Useful tip:**]

hitting letter "o" key will give you a panel view of the slides

---
class: middle

# Outline
1. Variance of Least squares Estimator 

???


---
class: inverse, center, middle
name: intro

# Variance of Least squares Estimator

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---

class: inverse, center, middle
name: intro

# Variance of Least squares Estimator

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>
---
class: middle

.content-box-red[**Important!**]

$$\mathbf{V}_{\hat{\beta}} = Var[\hat{\beta}|\mathbf{X}]=\mathbf{(X^{\prime}X)^{-1}X^{\prime}DX(X^{\prime}X)^{-1}}$$

, where $\mathbf{D}=Var[\mathbf{e|X}]=E[\mathbf{ee^{\prime}|X}]$

.content-box-green[**Exercise**]
+ Let's derive this formula. 


