---
title: "APEC8211: Recitation 3"
author: 
  - "Shunkei Kakimoto" 
header-includes:
   - \usepackage{mathtools}
   - \usepackage{color}
   - \usepackage{amsmath}
output:
  xaringan::moon_reader:
    self_contained: true
    css: ../xaringan-themer.css 
    lib_dir: libs
    nature: 
      highlightStyle: github
      highlightLines: true
---
class: middle

```{r, child = '../setup.rmd', cache = FALSE}
```

```{r, include = F, cache = FALSE}
# here::i_am("GitControlled/Recitation/1_Introduction/recitation1_slides.rmd")
```


```{r, include = F, cache = FALSE}
library(data.table)
library(ggplot2)
library(dplyr)
library(gganimate)
library(ggtext)
library(gifski)
library(gganimate)
```

```{r, include = F, eval=F, cache = FALSE}
httpgd::hgd()
httpgd::hgd_browse()
```

# Outline
+ 
+ CEF theorem
+ 
 

---
name: dist

class: middle

# Mean, Variance and Covariance:
.content-box-red[**Mean: E[X]**]
+ Definition: <span style='color:red'>The mean of $X$ is $E[X]$</span>
+ How to calculate it?:
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, dpi=36, echo=F, out.width = "50%", Cache=T}
x_left = 2
x_right = 10
x_center = (x_left+x_right)/2
x <- seq(x_left, x_right, length = 1000)
y <- dnorm(x, mean = x_center, sd = 1)
plot_data <- data.table(x = x, y = y)

ggplot(data = plot_data) +
  geom_line(aes(x = x, y = y), color = "black")+
  labs(y="Density", title="Density of X ~ N(6,1)") +
  geom_vline(xintercept=x_center, color="red") +
  annotate("text",
    x = 7, y = 0.01,
    label = "Mean E[X]=6",
    size = 3, color="red"
  ) +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Mean, Variance and Covariance:
.content-box-red[**Variance: Var[X]**]
+ Definition: <span style='color:red'>The variance of $X$ is $Var[X]=E[(X-E[X])^2]$</span>
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, dpi=36, echo=F, out.width = "50%"}
# /*===== Data generation =====*/
ls_res <- list()
ls_var <- c(1,4,9)

for (i in seq(1:length(ls_var))){
  var <- ls_var[i]
  x <- seq(x_left, x_right, length = 1000)
  y <- dnorm(x, mean = x_center, sd = sqrt(var))
  ls_res[[i]] <- data.table(var = var, x = x, y = y)
}

res_total <- rbindlist(ls_res)

# /*===== Visualization =====*/
ggplot(data = res_total) +
  geom_line(aes(x = x, y = y, color = interaction(var)))+
  labs(y = "Density", title = expression("Density of X ~ N(6," ~ sigma^2 ~ ") with various " ~ sigma^2)) +
  guides(color = guide_legend(title= expression(sigma^2 ~ "=")))+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


---
# Mean, Variance and Covariance:
.content-box-red[**Coariance: Cov[X, Y]**]
+ Definition: <span style='color:red'>The covariance between X and Y are</span>

$$
\color{red}{Cov(X, Y) = E[(X-E[X])((Y-E[Y]))]}
$$
+ How to calculate it?
  * for discrete $X$?
  * for continuous $X$?

+ .content-box-green[Visualization]

```{r, dpi=36,, echo=F, fig.dim = c(12, 3)}
# /*===== Data generation =====*/
ls_a <- c(-0.8, -0.5, 0, 0.5, 0.8)
n = 1000
ls_result <- list()

set.seed(123)
x <- rnorm(n, 3, sd=2)
z <- rnorm(n, 3, sd=2)

for (i in seq(1:length(ls_a))){  
  a <- ls_a[i]
  y <- a*x + abs(1-abs(a))*z
  ls_result[[i]] <- data.table(a = a, x = x, y = y)
}

results <- 
  rbindlist(ls_result) %>%
  .[,type := paste0("Cov(X,Y)=",round(cov(x,y), digits = 1)), by=a]

# /*===== Visualization =====*/
ggplot(data = results)+
  geom_point(aes(x = x, y = y), size = 0.5)+
  facet_wrap(~ factor(type, unique(results$type)), scale = "free", nrow = 1)+
  labs(x = "X", y = "Y", title = "Plots of random variables (X,Y) with different sizes of covariances")+
  theme_bw()+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

---
# Expectation, Variance and Covariance as operators
In Econometrics analysis, we often use $E[\,]$, $Var[\,]$, $Cov[\,]$ as operators (i.e., functions). The following are useful knowledge that you should know. 

---
name: exercise
class: middle

.content-box-red[**Expectation**] 

<span style="color:blue">Expectation is a linear operator.</span>

That is, for any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $E[a+bX]=a+bE[X]$

(2) $E[X+Y]=E[X]+E[Y]$

+ Note: Linearity of expectation only applies to the sum of random variables (random variables are in the linear function) 
<!-- + A function $g(x)$ called a linear operator if (1) $f(x+y)=f(x)+f(y)$, and (2) $g(cx)=cg(x)$ for all $x$ and constant $c$.  -->

.content-box-green[**Question**]
+ True or False: $E[X+X^2]=E[X]+E[X^2]$?
+ True or False: $E[XY]=E[X]E[Y]$

---
class: middle

.content-box-red[**Variance**]

<span style="color:blue">Variance is not a linear operator except for special case.</span>

For any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $Var[X]=E[X^2] - (E[X])^2 \quad (\text{Simply, another definition of } Var[X]$ )

(2) $Var[aX]=a^2Var[X]$

(3) $Var[a+bX]=b^2Var[X]$

(4) $Var[X+Y]=Var[X]+Var[Y] + Cov(X, Y)$

.content-box-green[**Question**]
+ In what condition does $Var[X+Y]=Var[X]+Var[Y]$ hold?


.content-box-green[**Exercise**]
+ Let's prove (1) and (2) 

---
class: middle

.content-box-red[**Visualization**]

For any constants $a$ and $b$ and any random variables $X$ and $Y$:

(1) $Cov[X, Y]=E[XY]-E[X]E[Y] \quad (\text{Simply, another definition of } Cov[X, Y]$)

(2) $Cov[aX, Y]=aCov[X,Y]$

(3) $Cov[X, a+bY]=bCov[X,Y]$

.content-box-green[**Question**]
+ In what condition does $Cov[X, Y]=0$ hold?

.content-box-green[**Exercise**]
+ Let's prove (1)-(3)

---



