---
title: \vspace{-2cm} Latex math notations and Example R codes with PDF output
author: "Shunkei Kakimoto"
output:
  pdf_document:
    number_sections: no
    toc: no
    toc_depth: 3
header-includes:
   - \usepackage{tabularx}    
   - \usepackage{verbatim}    
   - \usepackage{xcolor}    
   - \usepackage{bm}
geometry: margin=2cm
---
<!-- extra_dependencies: ["titling", "tabularx", "verbatim", "xcolor", "bm"] -->

<!-- new command -->
<!-- \newcommand{\gfcb}[1]{%
    \fcolorbox{white}{gray!10!}{\quad\strut #1\quad}
    } % gfcb := gray fcolorbox
\newcommand{\cop}[1]{%
    \gfcb{\texttt{\detokenize{#1}}} 
    \ensuremath{\quad\longrightarrow\quad #1}
    } % cop := code outpu -->


```{r setup, include=FALSE}
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE
)
```

<!-- # Objective
+ To show you can use Rmarkdown file with latex notation. 
+ To introduce some useful latex math notation to write math equations. 
+ For further basic information about how to use Rmarkdown, please refer [\textcolor{blue}{this}](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf). -->
 
# Section 1: Building blocks

\newsavebox{\pmat}
\savebox{\pmat}{$\displaystyle\begin{pmatrix}
1 & 2 & 3 \\
4 & 5 & 6 \\
\end{pmatrix}$} 

\newsavebox{\pmatcode}
\begin{SaveVerbatim}{pmatcode}
\begin{pmatrix}
  1 & 2 & 3 \\
  4 & 5 & 6 \\
\end{pmatrix}
\end{SaveVerbatim}

\newsavebox{\bmat}
\savebox{\bmat}{$\displaystyle\begin{bmatrix}
\beta_0 \\
\beta_1 \\
\beta_2 \\
\end{bmatrix}$}

\begin{SaveVerbatim}{bmatcode}
\begin{bmatrix}
  \beta_0 \\
  \beta_1 \\
  \beta_2 \\
\end{bmatrix}
\end{SaveVerbatim}

<!-- \begin{center} -->
\renewcommand\arraystretch{2}
<!-- \hspace*{-0.4in} -->
\begin{tabular}{|>{$\displaystyle}l<{$}|l|>{\footnotesize}>{\raggedright\arraybackslash}p{3in}|}
\omit\,\textbf{Goal}&\omit \textbf{\,How to get it}\hfill &\omit\textbf{\,Notes} \\
\hline
\alpha, \beta, \omega, \Omega &
\begin{BVerbatim}
\alpha, \beta, \omega, \Omega 
\end{BVerbatim}
& Spell names of Greek letters. \\
\hline
\bar X, \hat\beta, \tilde\beta &
\begin{BVerbatim}
\bar X, \hat\beta, \tilde\beta
\end{BVerbatim}
& Can't do stats without $\bar X$. Can't do econometrics without $\hat\beta$! \\
\hline 
\mathbf{X}, \boldsymbol{\beta} &
\begin{BVerbatim}
\mathbf{X}, \boldsymbol{\beta}
\end{BVerbatim}
& Bold math symbols. To make Latin-letters bold, you can use
\begin{BVerbatim} 
\boldsymbol
\end{BVerbatim}
 from "bm" package:
\begin{BVerbatim} 
\usepackage{bm} 
\end{BVerbatim}
\\
\hline 
\sigma^2 &
\begin{BVerbatim}
\sigma^2 
\end{BVerbatim}
& Superscripts. \\
\hline
\sim &
\begin{BVerbatim}
\sim
\end{BVerbatim}
& For example, $X \sim N(0, 1)$\\
\hline
\succ, \succsim&
\begin{BVerbatim}
\succ, \succsim
\end{BVerbatim}
& Preference relations in micro\\
\hline
\xrightarrow{p}, \xrightarrow{d}&
\begin{BVerbatim}
\xrightarrow{p}, \xrightarrow{d}
\end{BVerbatim}
& Convergence in probability, and convergence in distribution \\
\hline
\ge, \leq, >, <&
\begin{BVerbatim}
\ge, \leq, >, <
\end{BVerbatim}
& Inequality \\
\hline
X_i, \sigma_{ij} & \begin{BVerbatim}
X_i, \sigma_{ij} 
\end{BVerbatim}
& Subscripts. When a sub(super)script has more than one symbol, like the $ij$, braces around it are needed to say where the sub(super)script ends. \\
\hline
\sum_{i=1}^n X_i  & \begin{BVerbatim}
\sum_{i=1}^n X_i 
\end{BVerbatim}
& Summation.   \\
\hline
\prod_{i=1}^n X_i  & \begin{BVerbatim}
\prod_{i=1}^n X_i 
\end{BVerbatim}
& Products.\\
\hline
\int_{x=-\infty}^\infty x\,dx  & \begin{BVerbatim}
\int_{x=-\infty}^\infty x\,dx
\end{BVerbatim}
& Integrals. \\
\hline
\frac{1}{1-\beta}  & \begin{BVerbatim}
\frac{1}{1 - \beta}
\end{BVerbatim}
& Fractions. Braces enclose numerator and denominator. \\
\hline 
\frac{\partial f(x,y)}{\partial x} & \begin{BVerbatim}
\frac{\partial f(x,y)}{\partial x} 
\end{BVerbatim} 
& Partial derivative. \\
\hline
\log(x), \exp(x)  & \begin{BVerbatim}
\log(x), \exp(x) 
\end{BVerbatim}
& Named functions. Looks better if you don't omit the \verb|\|.\\
\hline
\sqrt{V+1}  & \begin{BVerbatim}
\sqrt{V+1} 
\end{BVerbatim}
& Square root.\\
\hline
\usebox{\pmat} &
 \BUseVerbatim[baseline=c]{pmatcode} & Matrices.\rule[-0.25in]{0pt}{0.6in}\\
\hline 
\usebox{\bmat} & \BUseVerbatim[baseline=c]{bmatcode} & Matrices.\rule[-0.33in]{0pt}{0.8in}\\
\hline
\end{tabular}
<!-- \enlargethispage{0.25in} -->

\newpage

# Section 2: Example math equations
## Example 1: Aligning equations with expression after equal sign
Syntax:
```{r, eval=F}
\begin{align*}
\hat{\beta} 
  &= \mathbf{(X^{\prime}X)^{-1} (X^{\prime}Y)} \\
  &= \mathbf{(X^{\prime}X)^{-1} (X^{\prime} (X \beta + \mathbf{e}))} \\
  &= \mathbf{(X^{\prime} X)^{-1} X^{\prime}{X} \beta 
   + (X^{\prime} X)^{-1} X^{\prime} e}\\ 
  &= \beta + \mathbf{(X^{\prime} X)^{-1} X^{\prime} e}
\end{align*}
````
becomes:
\begin{align*}
\hat{\beta} 
  &= \mathbf{(X^{\prime}X)^{-1} (X^{\prime}Y)} \\
  &= \mathbf{(X^{\prime}X)^{-1} (X^{\prime} (X \beta + \mathbf{e}))} \\
  &= \mathbf{(X^{\prime} X)^{-1} X^{\prime}{X} \beta 
   + (X^{\prime} X)^{-1} X^{\prime} e}\\ 
  &= \beta + \mathbf{(X^{\prime} X)^{-1} X^{\prime} e}
\end{align*}



\

## Example 2: Writing equations within a curly brace
Syntax:
```{r, eval=F}
\begin{equation*}
  F(x) =
    \begin{cases}
      0 &  \quad x < 0 \\
      x &  \quad 0 \ge x \ge 1\\
      1 &  \quad x > 1
    \end{cases}       
\end{equation*}
```
becomes:
\begin{equation*}
  F(x) =
    \begin{cases}
      0 & x < 0 \\
      x & 0 \ge x \ge 1\\
      1 & x > 1
    \end{cases}   
\end{equation*}

\newpage

## Example 3: Including comments within equations
Syntax:
```{r, eval=F}
\begin{equation*}
  Z_n =
    \begin{cases}
      -n & \text{with probability $1/n$}\\
      0 &  \text{with probability $1-2/n$}\\
      n &  \text{with probability $1/n$}\\
    \end{cases}       
\end{equation*}
```
becomes:
\begin{equation*}
  Z_n =
    \begin{cases}
      -n & \text{with probability $1/n$}\\
      0 &  \text{with probability $1-2/n$}\\
      n &  \text{with probability $1/n$}\\
    \end{cases}       
\end{equation*}

Sintax:
```{r, eval=F}
\begin{align*}
\overline{X}_n = 
  \frac{1}{n}\sum{i=1}^{n}X_i \xrightarrow{p} E[X] 
  \quad \text{ as } n \rightarrow \infty
\end{align*}
```
becomes:
\begin{align*}
\overline{X}_n = \frac{1}{n}\sum_{i=1}^{n}X_i \xrightarrow{p} E[X] \quad \text{ as } n \rightarrow \infty
\end{align*}

\

## Example 4: Maximization problem
Syntax:
```{r, eval=F}
\begin{equation*}
\begin{aligned}
\max_{x_1, x_2} \quad & u(x_1, x_2) = \frac{1}{2}log x_1 + \frac{1}{2} log x_2 \\
\textrm{s.t.} \quad & w=p_1 x_1 + p_2 x_2 \\
\end{aligned}
\end{equation*}
```
becomes:
\begin{equation*}
\begin{aligned}
\max_{x_1, x_2} \quad & u(x_1, x_2) = \frac{1}{2}log x_1 + \frac{1}{2} log x_2 \\
\textrm{s.t.} \quad & w=p_1 x_1 + p_2 x_2 \\
\end{aligned}
\end{equation*}

\newpage 

# Section 3: Write R codes

```{r}
# === Load Packages === #
library(data.table)
library(ggplot2)

# === Data === #
#built-in dataset in R
data(iris) 
# Convert the data into data.table
setDT(iris)
# Filter for the species "setosa" 
virginica <- iris[Species == "virginica",]
```

```{r, out.width="80%"}
# === Visualization === #
ggplot(virginica, aes(x=Sepal.Length, y=Petal.Length))+
  geom_point()+
  geom_smooth(method = lm, se = FALSE)+
  labs(title = "Relationship between Sepal.Length and Petal.Length")+
  theme(plot.title = element_text(hjust = 0.5))
```


# Section 4: Insert a picture to a document
Below is an example R chunk code to insert a picture. 
````markdown
`r ''````{r, fig.cap='A caption', out.width = '80%'}
knitr::include_graphics("path to the file")
```
````